Biodiversity changes are linked to changes in food-web structure and redundancy in riverine ecosystems

### 2025-02-03 paper/appendix.Rmd
Manuscript update before Maud & Colin comments

---
title: Appendix
subtitle:
author: Alain Danet, Loubna El Madouri, Maud Mouchet, Elisa Thébault and Colin Fontaine
date: \today
output:
  bookdown::pdf_document2:
    fig_caption: true
    keep_tex: true
    toc: false
fontsize: 10pt
header-includes:
   - \usepackage{natbib}
   - \newcommand{\beginsupplement}{\renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
   - \usepackage{float}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
   - \renewcommand{\arraystretch}{1}
geometry: margin=2cm
bibliography: references.bib
number_sections: false
---

>>>>

---
title: "Biodiversity changes are linked to changes in food-web structure and redundancy in riverine ecosystems"
subtitle: "Alain Danet (a,b), Loubna El Madouri (a,c), Andrew P. Beckerman (b),
Maud Mouchet (a), Colin Fontaine (a) and Elisa Thébault (d)"
date: Supplementary figures and tables
output:
  bookdown::pdf_document2:
    fig_caption: true
    keep_tex: true
    toc: false
fontsize: 10pt
header-includes:
   - \usepackage{natbib}
   - \newcommand{\beginsupplement}{\renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
   - \usepackage{float}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
   - \renewcommand{\arraystretch}{1}
geometry: margin=2cm
bibliography: references.bib
number_sections: false
---

### 2025-02-03 paper/appendixB.Rmd
Manuscript update before Maud & Colin comments

---
title: Appendix B - Dominator tree and redundancy
subtitle:
author:
date: \today
output:
  bookdown::pdf_document2:
    fig_caption: true
    keep_tex: true
    toc: false
fontsize: 10pt
header-includes:
   - \newcommand{\beginsupplement}{\renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
   - \usepackage{float}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
   - \renewcommand{\arraystretch}{1}
geometry: margin=2cm
bibliography: references.bib
number_sections: false
---

@@

Because species are interdependant, species loss can have cascading effects and
triger secondary extinctions [@dunne_network_2002]. So assessing the robustness
of ecological networks to species losses has been a big question in Ecology.
Previous research showed that food-web structure, such as connectance, increased
robustness [@dunne_network_2002]. The idea is that more connected food-web will
be more stable because they contain a large number of pathways convey biomass
through the food-web. However, connectance as some limitations as a robustness
metric has some limitations, because each trophic link does not contribute
equally to the robustness of an ecological network. Trophic links can be as
functional or redundant, depending on their role in transmitting biomass through
the food-web. Functional links are the set of links that is required to connect
every species to the basal resource of the food-web, all of the other being
redundant. Only functional links affect directly robustness, as their removal
will trigger secondary extinctions. The removal of redundant links in turn do
not trigger secondary extinctions, but makes the food-webs more fragile to
further extinctions [@allesina_functional_2009].

The identification of functional and redundant links is based on the use of
denomitor trees [@bodini_using_2009; @allesina_who_2004;
@allesina_functional_2009]. Dominator trees allows to identify the set of
trophic links that are essential to connect species to basal resources, i.e.
functional links, and the redundant links. Pioneer papers explained the
algorithm to compute the set of dominator trees and I refer to them for further
information [@bodini_using_2009; @allesina_who_2004; @allesina_functional_2009]
and examples. Here, we aim to show that while implementing this algorithm could
seem somewhat intimitating at the time, this algorithm is now widely
implemented. In R, it is available in `igraph::dominator_tree()`. Below we show
how functional and redundant links can be computed easily from R by replicating
three examples presented in previous studies.

@@

We used this function to reproduce the three examples (Fig. \@ref(fig:examples)),
which were displayed in the Fig.1 and Appendix 1 of @bodini_using_2009, and
in the Fig.1 of @allesina_functional_2009. The table \@ref(tab:redun) displays
the number and the proportion of functional and redundant links.

@@

(ref:example-cap) Reproduced examples of categorising functional links in
food-webs using `igraph::dominator_tree()` R function. Left columns (a, c, e)
displays the complete food-web while right column displays food-web with the
functional links only (b, d, f). (a,b): reproduction of Fig. 1 of
@bodini_using_2009, (c, d): reproduction of appendix figures of @bodini_using_2009,
(e, f): reproduction of fig. 1 of @allesina_functional_2009. Note that (e, f) is
a bit different from the original, as noted by the authors, they could have
removed further redundant links (i.e. $2 \to 5$, or $3 \to 5$).

>>>>

---
title: "Biodiversity changes are linked to changes in food-web structure and redundancy in riverine ecosystems"
subtitle: "Alain Danet (a,b), Loubna El Madouri (a,c), Andrew P. Beckerman (b),
Maud Mouchet (a), Colin Fontaine (a) and Elisa Thébault (d)"
date: Appendix B - Computation of redundancy of trophic pathways in food webs
output:
  bookdown::pdf_document2:
    fig_caption: true
    keep_tex: true
    toc: false
fontsize: 10pt
header-includes:
   - \newcommand{\beginsupplement}{\renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
   - \usepackage{float}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
   - \renewcommand{\arraystretch}{1}
geometry: margin=2cm
bibliography: references.bib
number_sections: false
---

@@

Because species biomass are involved in trophic interactions, species losses can
have cascading effects and trigger secondary extinctions [@dunne_network_2002].
In the context of biodiversity crisis, assessing the robustness of ecological
networks to species losses has been then a big question in Ecology. Previous
research showed that food-web structure, such as connectance, increased
robustness [@dunne_network_2002]. The idea is that more connected food-web will
be more stable because they contain a large number of pathways convey biomass
through the food-web. However, connectance as some limitations as a robustness
metric, because each trophic link does not contribute equally to the robustness
of an ecological network [@allesina_who_2004]. Trophic links can be categorised as
functional or redundant, depending on their role in transmitting biomass through
the food-web. Functional links are the set of links that is required to connect
every species to the basal resource of the food-web, all of the other being
redundant. Only functional links affect directly robustness, as their removal
will trigger secondary extinctions. The removal of redundant links in turn do
not directly trigger secondary extinctions, but makes the food-webs more fragile
to further extinctions [@allesina_functional_2009].

The identification of functional and redundant links is based on the use of
denominator trees [@bodini_using_2009; @allesina_who_2004;
@allesina_functional_2009]. Dominator trees allows to identify the set of
trophic links that are essential to connect all the consumers to the basal resource, i.e.
the functional links, and the redundant links, i.e. the redundant trophic pathways
from basal resource to consumers. Pioneer papers explained the algorithm to
compute the set of dominator trees and we refer to them for further
information and examples [@bodini_using_2009; @allesina_who_2004;
@allesina_functional_2009]. Here, we aim to show that while implementing this
analysis could be difficult at the time, this algorithm is now widely
available. In R, it is available in `igraph::dominator_tree()`. Below we show
how the number of functional and redundant links can be  easily computed from R
by replicating three examples presented in previous studies.

@@

We used this function to reproduce the three examples (Fig. \@ref(fig:examples)),
which were displayed in the Fig. 1 and Appendix 1 of @bodini_using_2009, and
in the Fig. 1 of @allesina_functional_2009. The table \@ref(tab:redun) displays
the number and the proportion of functional and redundant links from those
examples.

@@

(ref:example-cap) Reproduced examples of categorising functional links in
food webs using `igraph::dominator_tree()` R function. Left columns (a, c, e)
displays the complete food-web while right column displays food-web with the
functional links only (b, d, f). (a,b): reproduction of Fig. 1 of
@bodini_using_2009, (c, d): reproduction of appendix figures of @bodini_using_2009,
(e, f): reproduction of fig. 1 of @allesina_functional_2009. Note that (e, f) is
a bit different from the original, as noted by the authors, they could have
removed further redundant links (i.e. $2 \to 5$, or $3 \to 5$).
